---
title: "R Notebook"
output:
  pdf_document: default
  html_document:
    df_print: paged
---



```{r}
library(tidyverse)
```
mtcars is a dataset in which each observation is a car model and has 11 variables. The data is from 1974.
```{r}
# call built-in data mtcars.
data(mtcars)

# Select only car models where mpg<20
mtcars_mpg2 <- mtcars[mtcars$mpg < 20,]


# Reduce the variables to mpg, cyl, disp, hp, gears
mtcars_mpg2 <- mtcars_mpg2[, c(1,2,3,4,10)]

# read the R file hand_functions.R so that it can be used
# notice that with echo = TRUE 
source(file = "hand_functions.R", echo = TRUE)

# Now use the function from hand_functions.R to produce summary statistics

sp_out <- sum_special(mtcars_mpg2)
sp_out


# library(esquisse)
# 
# esquisser(data = mtcars_mpg2, viewer = "browser")
```

  The first plot shows mpg vs. engine displacement. The plot shows that mpg decreases with displacement. Displacement is the combined swept volume of the pistons inside the cylinders of an engine and directly affects power output and fuel efficiency.
```{r}
ggplot(mtcars_mpg2) +
  aes(x = disp, y = mpg) +
  geom_point(shape = "bullet", size = 4L, colour = "#B22222") + #scatterplot
  geom_smooth(span = 0.5) + #line plot, span controls the amount of smoothing for the default loess smoother
  theme_minimal() #theme, how the plot looks
```

  The second plot shows that cars with 6 cylinders have higher mpg than cars with 8 cylinders.
```{r}
# note that this boxplot cannot be made with esquisse() unless
# `cyl` is changed to a factor using `as.factor(mtcars_mpg2$cyl)`.
# I verified this with this code:
# mtcars_mpg3 <- mtcars_mpg2
# mtcars_mpg3$cyl <- as.factor(mtcars_mpg3$cyl)
# esquisser(data = mtcars_mpg3)

ggplot(mtcars_mpg2, aes(x=as.factor(cyl), y=mpg)) + #`as.factor` changes `cyl` to a factor or category
  geom_boxplot(fill="slateblue", alpha=0.2) + #`fill` fills the boxes with color, `alpha` changes transparency, effectively lightening the color
  xlab("cyl")
```

